### 2018-05-10 thesis_report/Thesis_Model_Improvements.tex
minor fixes.
Model Improvements v1

By applying this new labeling, we refine our previous datasets $\CC_{strict}$, $\CC_{loose}$ and $\CC_{real}$ as shown in Table \ref{tab:dataset_sizes_new}.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		\toprule
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\midrule
		Trojan			& 0 & 0 & 0\\
		Virus			& 0 & 0 & 0\\
		Worm			& 0 & 0 & 0\\
		Spyware			& 0 & 0 & 0\\
		Ransom			& 0 & 0 & 0\\
		Other			& 0 & 0 & 0\\
		\midrule
		Total Malware	& 0 & 0 & 0\\
		Total Goodware	& 0 & 0 & 0\\
		\midrule
		\midrule
		Total			& 0 & 0 & 0\\
		\bottomrule
	\end{tabular}
	\caption{New sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_new}
\end{table}

@@

With that in mind, we decide to transform the values into a $[0,1]$ range.
We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large values fall into the same bin.

@@

For our third and last type of dynamic information, we resort to Cuckoo~\cite{tool:cuckoo} custom signatures.
These signatures are built from certain activities that Cuckoo deems malicious or suspicious.
For example, if a sample allocates memory and then makes it executable, it might suggest some sort of packing or obfuscation.

Although some signatures are defined by a sequence of library calls, which could also be attained by creating n-grams of our previously defined feature, using the already defined Cuckoo signatures saves training time.
As an example, creating bi-grams of library calls, given there are 144 unique calls, would yield over 20,000 possible features, hence being unpractical.

@@

Having these new features, we create a new set with all features (static, category and library calls and signatures) $\FF_{dynamic}$.
We joined these samples $\FF_{dynamic}$ to the labeled samples $\CC_{real}$, obtaining a total of \miss{size} labeled samples with dynamic imports $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$.

% To reduce the number of features into those that are more representative, we apply, as in other features, a variance threshold, removing signatures that do not vary in more than 80\% of the samples.
% By doing so, we reduce the total number of signatures to

@@

In the previous sections we described all our modifications to the original model $\LR$: a two layer approach that splits malware into subclasses and the addition of three dynamic features.
We now present the results of our new model $\mathcal{E}$, validated using the same methodology as described in \ref{section:evaluation}.
Specifically we test the model using the baseline cross-validation methodology, followed by our three temporally consistent scenarios: \textit{Past-to-Present}, \textit{Present-to-Past} and \textit{Temporal Window}.
We test each methodology using the three different scenarios: $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.

@@

Starting with \textit{laboratory conditions}, we apply the cross-validation evaluation to model $\mathcal{E}$ with the labeled dataset $\CC_{strict}$ and features $\DD_{dynamic}$, providing scenario $\mathcal{S}_{\strict}$.
For this scenario, we obtain an \gls{auroc} of \miss{score}, as presented by the red line in Figure \ref{fig:improved_xval}.
These provide the highest results, provided by factors like a small and reliable dataset, and the use of cross-validation.

Relaxing to more \textit{real-world conditions}, under the form of a less reliable ground truth, we test the datasets $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{dynamic}$.
As shown in Figure \ref{fig:improved_xval}, the score under \gls{auroc} decreases to \miss{socre} for $\mathcal{S}_{\loose}$ (blue line) and \miss{score} for $\mathcal{S}_{\real}$ (green line).

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/dia_slidingwindow.png}
	\caption{Improved model results with dynamic features in laboratory conditions.}
	\label{fig:improved_xval}
\end{figure}

@@

>>>>

The following Table \ref{tab:dataset_sizes_new} discriminates the amount of each malware class.
Note that the classes are not exclusive, as such the represented numbers show how many samples, out of each dataset $\CC_{strict}$, $\CC_{loose}$ and $\CC_{real}$, are classified by at least one vendor as belonging to any of the defined classes.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		\toprule
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\midrule
		Trojan			& 97,054 & 44,329 & 24,176\\
		Other			& 49,443 & 24,126 & 14,750\\
		Worm			& 24,554 & 14,837 & 9,381\\
		Virus			& 21,055 & 12,531 & 6,899\\
		Spyware			& 20,724 & 10,172 & 5,955\\
		Ransom			& 7,761 & 1,924 & 1,160\\
		\bottomrule
	\end{tabular}
	\caption{Samples belonging to each of the 6 malware classes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_new}
\end{table}

One can see that the \textit{trojan} class is present in almost all samples, whereas \textit{ransom} is the least common class.
More detailed statistics regarding each dataset is presented in Appendix \ref{chapter:appendixClasses}, containing information like the average weight for each class and the minimum and maximum values.

@@

With that in mind, we decide to transform the values to follow a normal distribution.
We do this by using scikit-learn's~\cite{tool:sklearn} \textit{QuantileTransformer} with a normal distribution, which splits the possible values into bins such that the resulting distribution is of type \textit{Gaussian} with a mean of 0.
This way we have a greater number of bins around the mean, allowing for better discrimination, whereas very large values fall into the same bin.

@@

For our third and last type of dynamic information, we resort to Cuckoo's~\cite{tool:cuckoo} custom signatures.
These signatures are built from certain activities that Cuckoo deems malicious or suspicious.
For example, if a sample allocates memory and then makes it executable, it might suggest some sort of packing or obfuscation.

Although some signatures are defined by a sequence of library calls, which could also be attained by creating \textit{n-grams} of our previously defined feature, using the already defined Cuckoo signatures saves training time.
As an example, creating bi-grams of library calls, given there are 144 unique calls, would yield over 20,000 possible features, which becomes highly unpractical when considering a higher degree of n-grams.

@@

We joined these features $\FF_{dynamic}$ to the labeled samples $\CC_{real}$, obtaining a total of 122,633 labeled samples with the new features $\DD_{dynamic} = \CC_{real} \cap \FF_{dynamic}$.

Given there is a lower amount of available samples, we provide in Table \ref{tab:dataset_sizes_improved} the new sizes for $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$, which take into account the new features $\FF_{dynamic}$.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{lccc}
		\toprule
		Dataset			& $\CC_{real}$ & $\CC_{loose}$ & $\CC_{strict}$	\\
		\midrule
		Malware			& 94,248 & 42,911 & 23,437\\
		Goodware		& 28,385 & 1,741 & 1,741\\
		\midrule
		Total			& 122,633 & 44,652 & 25,178\\
		\bottomrule
	\end{tabular}
	\caption{New sizes for datasets $\CC_{real}$, $\CC_{loose}$ and $\CC_{strict}$.}
	\label{tab:dataset_sizes_improved}
\end{table}

@@

In the previous sections we described all our modifications to the original model $\LR$: a two layer approach that splits malware into subclasses and the addition of three dynamic features.
We now present the results of our new model $\mathcal{E}$, validated using the same methodology as described in \ref{section:evaluation}.

Specifically we test the model using the baseline cross-validation methodology, followed by our three temporally consistent scenarios: \textit{Past-to-Present}, \textit{Present-to-Past} and \textit{Temporal Window}.
We test each methodology using the three different scenarios: $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.

@@

Starting with \textit{laboratory conditions}, we apply the cross-validation evaluation to model $\mathcal{E}$ with the labeled dataset $\CC_{strict}$ and features $\DD_{dynamic}$, providing scenario $\mathcal{S}_{\strict}$.
For this scenario, we obtain an \gls{auroc} of 98\%, as presented by the red curve in Figure \ref{fig:improved_xval}.
These provide the highest results, provided by factors like a small and reliable dataset, more features and the use of cross-validation.

Relaxing to more \textit{real-world conditions}, under the form of a less reliable ground truth, we test the datasets $\CC_{loose}$ and $\CC_{real}$ on features $\DD_{dynamic}$.
As shown in Figure \ref{fig:improved_xval}, the score under \gls{auroc} is 98\% for $\mathcal{S}_{\loose}$ (blue curve) and 95\% for $\mathcal{S}_{\real}$ (green curve).

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/xval_improved.png}
	\caption{Multi layer results for dynamic features in laboratory conditions.}
	\label{fig:improved_xval}
\end{figure}

@@

As with previous results, one notes that from $\mathcal{S}_{\strict}$ to $\mathcal{S}_{\loose}$ the results are affected at all, when the change between the scenarios is merely in the number of malware samples.
Between $\mathcal{S}_{\loose}$ to $\mathcal{S}_{\real}$ we again note the already seen pattern, the score is lowest when using the most realistic dataset.

The comparison between scenarios does not yield any new information from what was seen in Section \ref{section:single_layer_results}.
What is more interesting is that the absolute values are boosted in all scenarios, which show indeed that using the multi-layer approach with dynamic features improve the model's results.

Having applied the same cross-validation to our modified model, obtaining interesting results, we now go over to test how our temporal based methodologies are affected.

We start again with \textit{Past-to-Present} validation to each scenario $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$.
Figure \ref{fig:pastpresent_modified} shows the \gls{auroc} at every iteration (\ie\ fold) for each scenario: 96\% for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$, and 92\% $\mathcal{S}_{\real}$.

\begin{figure}[!htb]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/pastpresent_improved.png}
	\caption[Multi layer results for dynamic features in \textit{Past-to-Present}.]{\gls{auroc} for each iteration of the \textit{Past-to-Present} evaluation. Folds order consistent with temporal order (\ie\ fold 0 contains older samples than
		fold 1)}
	\label{fig:pastpresent_modified}
\end{figure}

When comparing the average \gls{auroc} between cross-validation and \textit{Past-to-Present} validation, we note that both $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ decrease 2\%, while $\mathcal{S}_{\real}$ decreases 3\%.
This decrease is not a surprise, given the temporal consistency enforcement between samples.

The results are consistent with was previously seen in Section \ref{section:single_layer_results}, with the added factor that the absolute values are higher, and the tendency to increase is more present as we move forward in time, close to the validation set.

Following the previous evaluation order, we now present the results using our \textit{Present-to-Past} validation methodology.
In Figure \ref{fig:presentpast_improved} we present the \gls{auroc} for each iteration, where higher folds represent older samples. Here we see values of 97\% for $\mathcal{S}_{\strict}$, 98\% for $\mathcal{S}_{\loose}$ and 96\% $\mathcal{S}_{\real}$.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/presentpast_improved.png}
	\caption[Multi layer results for dynamic features in \textit{Present-to-Past}.]{\gls{auroc} for each iteration of the \textit{Present-to-Past} evaluation. Folds order is the inverse of temporal order (\ie\ fold 0 contains newer samples than fold 1)}
	\label{fig:presentpast_improved}
\end{figure}

When comparing to cross-validation, we note $\mathcal{S}_{\strict}$ is affected as expected, whereas $\mathcal{S}_{\loose}$ is not.
The fact that the datasets vary in size and so the amount of goodware and malware used for testing in each may vary, can justify how the $\mathcal{S}_{\loose}$ is not affected, whereas $\mathcal{S}_{\strict}$ is.

$\mathcal{S}_{\real}$ seems to have a higher value than in cross-validation, but it is rounded, which slightly inflates the value.
In practice both cross-validation and \textit{Present-to-Past} are identical for $\mathcal{S}_{\real}$.

In these results, the noticeable increase in the first 3 folds (0, 1 and 2) for $\mathcal{S}_{\strict}$ and $\mathcal{S}_{\loose}$ goes even more in favor with our argument that samples closer to the validation set benefit the model.
More so as the \gls{auroc} stabilizes from those folds on.
This effect is not as accentuated for $\mathcal{S}_{\real}$, although using more and older folds do not provide significantly better results.

Finally, we retest how a reduced training set behaves in our scenarios by using our \textit{Temporal Window} methodology.
As previously mentioned, the first 3 folds seem to provide enough information to obtain good results, hence we apply the same sliding window size of $w=3$ as in Section \ref{section:single_layer_results}.
In summary, we selected 30\% of each dataset for training and 10\% for validation.
Starting at the oldest fold, we apply this window and slide it by one fold at each iteration.
Figure \ref{fig:slidingwindow_modified} shows how all our scenarios $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ and $\mathcal{S}_{\real}$ score the same \gls{auroc} of 94\%, although with different curves.

For $\mathcal{S}_{\strict}$, $\mathcal{S}_{\loose}$ the score is equally and negatively affected by 4\%.
The jagged curve on both scenarios indicate that slight changes on the \gls{fpr} threshold have significant impact on the \gls{tpr}, this may be caused by the smaller dataset size, which in turn creates uneven folds for malware and goodware.

\begin{figure}[!h]
	\centering
	\includegraphics[width=0.8\textwidth]{Figures/slidingwindow_improved.png}
	\caption[Multi layer results for dynamic features in \textit{Temporal Window}.]{\gls{auroc} for our three scenarios, under the \textit{Temporal Window} methodology.}
	\label{fig:slidingwindow_modified}
\end{figure}

For $\mathcal{S}_{\real}$, the fact that it only loses 1\% when compared to cross-validation shows that indeed our argument about reducing the training set size is sustained.
Its curve is much smoother when compared to the previous scenarios, as the dataset is bigger and more even.

As a closing remark, we reaffirm on how it is possible to yield reasonable results when considering temporal consistent samples, specifically in this improved model, which not only uses more features, but has the ability to provide more information regarding a malware sample, as it can discriminate malware classes due to its multi layer design.
The improved layer results are summarized in Table \ref{tab:multilayer_results}.

\begin{table}[!htb]
	\renewcommand{\arraystretch}{1.2} % more space between rows
	\centering
	\begin{tabular}{ccccc}
		\toprule
		AUROC & $\DS_\strict$ & $\DS_\loose$ & $\DS_\real$ & Train/Test \%\\
		\midrule
		Cross-Validation & 0.98 & 0.98 & 0.95 & 90 / 10\\
		Past-to-Present & 0.96 & 0.96 & 0.92 & 10 to 90 / 10\\
		Present-to-Past & 0.97 & 0.98 & 0.96 & 10 to 90 / 10\\
		Sliding-Window & 0.94 & 0.94 & 0.94 & 30 / 10\\
		\bottomrule
	\end{tabular}
	\caption{Multi layer results summary.}
	\label{tab:multilayer_results}
\end{table}

Having studied how our modifications improved the proposed model, the following chapter goes on to provide practical applications for our work.
